---
title: "Fit VBGE models to back-calculated length-at-age"
author: "Max Lindmark, Jan Ohlberger, Anna G책rdmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    embed-resources: true
    fig-width: 12
    fig-height: 10
editor: source
execute: 
  echo: true
  eval: true
  cache: true
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```

## Load libraries

```{r}
#| cache=FALSE
pkgs <- c("tidyverse", "tidylog", "rTPC", "nls.multstart", "broom", "RColorBrewer", "viridis", "minpack.lm", "patchwork", "ggtext")

## minpack.lm needed if using nlsLM()
if(length(setdiff(pkgs, rownames(installed.packages()))) > 0){

    install.packages(setdiff(pkgs, rownames(installed.packages())), dependencies = T)
  
  }

invisible(lapply(pkgs, library,character.only = T))

# devtools::install_github("seananderson/ggsidekick") ## not on CRAN 
library(ggsidekick)
theme_set(theme_sleek())

# Check the temperature model script! This is the order based on mean temperature, which makes sense for the sharpe schoolfield plot, and therefore we might keep it across plots
order <- data.frame(area = c("SI_HA", "BT", "TH", "VN", "SI_EK", "FM", "JM", "BS", "MU", "FB", "HO", "RA")) 

nareas <- length(unique(order$area)) + 1
colors <- colorRampPalette(brewer.pal(name = "RdYlBu", n = 10))(nareas)[-c(7)]

# Load functions
home <- here::here()
fxn <- list.files(paste0(home, "/R/functions"))
invisible(sapply(FUN = source, paste0(home, "/R/functions/", fxn)))
```

Load cache

```{r}
# qwraps2::lazyload_cache_dir(path = paste0(home, "/R/analyze-data/02-fit-vbge_cache/html"))
```

## Read and trim data

```{r}
d <- #read_csv(paste0(home, "/data/for-analysis/dat.csv")) |> 
  read_csv("https://raw.githubusercontent.com/maxlindmark/perch-growth/master/data/for-analysis/dat.csv") |> 
  filter(age_ring == "Y") # use only length-at-age by filtering on age_ring

# Sample size by area and cohort
ns <- d |> 
  group_by(cohort, area) |> 
  summarise(n = n())

# Minimum number of observations per area and cohort
d$area_cohort <- as.factor(paste(d$area, d$cohort))

d <- d |>
  group_by(area_cohort) |> 
  filter(n() > 100)

# Minimum number of observations per area, cohort, and age
d$area_cohort_age <- as.factor(paste(d$area, d$cohort, d$age_bc))

d <- d |>
  group_by(area_cohort_age) |> 
  filter(n() > 10)

# Minimum number of cohorts in a given area
cnt <- d |>
  group_by(area) |>
  summarise(n = n_distinct(cohort)) |>
  filter(n >= 10)

d <- d[d$area %in% cnt$area, ]

# Plot cleaned data
ggplot(d, aes(age_bc, length_mm)) +
  geom_jitter(size = 0.1, height = 0, alpha = 0.1) +
  scale_x_continuous(breaks = seq(20)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 15)) +
  labs(x = "Age", y = "Length (mm)") +
  guides(color = "none") + 
  facet_wrap(~area, scale = "free_x")

# Longitude and latitude attributes for each area
area <- c("BS", "BT", "FB", "FM", "HO", "JM", "MU", "RA", "SI_EK", "SI_HA", "TH", "VN")
nareas <- length(area)
lat <- c(60, 60.4, 60.3, 60.5, 63.7, 58, 59, 65.9, 57.3, 57.4, 56.1, 57.5)
lon <- c(21.5, 18.1, 19.5, 18, 20.9, 16.8, 18.1, 22.3, 16.6, 16.7, 15.9, 16.9)
area_attr <- data.frame(cbind(area = area, lat = lat, lon = lon)) |>
  mutate_at(c("lat","lon"), as.numeric)
```

## Fit VBGE models

```{r}
# Get individual growth parameters (functions: VBGF/Gompertz,nls_out,fit_nls)
IVBG <- d |> 
  group_by(ID) |> 
  summarize(k = nls_out(fit_nls(length_mm, age_bc, min_nage = 4, model = "VBGF"))$k,
            k_se = nls_out(fit_nls(length_mm, age_bc, min_nage = 4, model = "VBGF"))$k_se,
            linf = nls_out(fit_nls(length_mm, age_bc, min_nage = 4, model = "VBGF"))$linf,
            linf_se = nls_out(fit_nls(length_mm, age_bc, min_nage = 4, model = "VBGF"))$linf_se)
```

## Inspect predictions

```{r}
#| include: false 
#| eval: false 

na_ids <- IVBG |>
  mutate(na = ifelse(is.na(k_se) == TRUE, "na", "not na")) |>
  select(ID, na)

test <- d |>
  ungroup() |>
  left_join(na_ids)

test |>
  mutate(age_group = ifelse(age_catch >= 4, "4 or older", "1-3")) |>
  ggplot(aes(age_catch, fill = age_group)) +
  geom_histogram() +
  facet_wrap(~na, ncol = 1)

# There are some NA's that *are* the right age, and no NAs below the correct age
# They just don't seem to converge!
test2 <- test |>
  ungroup() |>
  filter(na == "na") |>
  filter(age_catch > 4) |> # age 4 was the min_age argument
  arrange(ID) |>
  as.data.frame()

test3 <- d |> filter(ID == "1970.211.JM.female.5.K010")

plot(test3$age_bc, test3$length_mm)

nls_out(fit_nls(test3$length_mm, test3$age_bc, min_nage = 4, model = "VBGF"))
```

```{r}
IVBG <- IVBG |> drop_na(k) # The NAs are because the model didn't converge or because they were below the threshold age

IVBG <- IVBG |>
  mutate(k_lwr = k - 1.96*k_se,
         k_upr = k + 1.96*k_se,
         linf_lwr = linf - 1.96*linf_se,
         linf_upr = linf + 1.96*linf_se,
         row_id = row_number())

# Plot all K's
IVBG |>
  #filter(row_id < 5000) |>
  ggplot(aes(row_id, k, ymin = k_lwr, ymax = k_upr)) +
  geom_point(alpha = 0.5) +
  geom_errorbar(alpha = 0.5) +
  NULL

# Plot all L_inf's
IVBG |>
  #filter(row_id < 5000) |>
  ggplot(aes(row_id, linf, ymin = linf_lwr, ymax = linf_upr)) +
  geom_point(alpha = 0.5) +
  geom_errorbar(alpha = 0.5) +
  NULL

# We can also consider removing the upper 95% quantile of standard errors (not quantile of K directly)
IVBG |>
  tidylog::mutate(keep = ifelse(k > quantile(k_se, probs = 0.95), "N", "Y")) |>
  #filter(row_id < 10000) |>
  ggplot(aes(row_id, k, ymin = k_lwr, ymax = k_upr, color = keep)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~keep, ncol = 1) +
  geom_errorbar(alpha = 0.5) +
  NULL

# Add back cohort and area variables
IVBG <- IVBG |> 
  left_join(d |> select(ID, area_cohort)) |> 
  separate(area_cohort, into = c("area", "cohort"), remove = TRUE, sep = " ") |> 
  mutate(cohort = as.numeric(cohort))

# Summarise and save for sample size
sample_size <- IVBG |>
  group_by(area) |> 
  summarise(n_cohort = length(unique(cohort)),
            min_cohort = min(cohort),
            max_cohort = min(cohort),
            n_individuals = length(unique(ID)),
            n_data_points = n())

sample_size

sample_size |>
  ungroup() |>
  summarise(sum_ind = sum(n_individuals), sum_n = sum(n_data_points))

write_csv(sample_size, paste0(home, "/output/sample_size.csv"))

# Compare how the means and quantiles differ depending on this filtering
IVBG_filter <- IVBG |>
  tidylog::drop_na(k_se) |>
  tidylog::filter(k_se < quantile(k_se, probs = 0.95))

# Summarize growth coefficients by cohort and area
VBG <- IVBG |>
  group_by(cohort, area) |>
  summarize(k_mean = mean(k, na.rm = T),
            k_median = quantile(k, prob = 0.5, na.rm = T),
            linf_median = quantile(linf, prob = 0.5, na.rm = T),
            k_lwr = quantile(k, prob = 0.05, na.rm = T),
            k_upr = quantile(k, prob = 0.95, na.rm = T)) |> 
  ungroup()

VBG_filter <- IVBG_filter |>
  group_by(cohort, area) |>
  summarize(k_mean = mean(k, na.rm = T),
            k_median = quantile(k, prob = 0.5, na.rm = T),
            k_lwr = quantile(k, prob = 0.05, na.rm = T),
            k_upr = quantile(k, prob = 0.95, na.rm = T)) |> 
  ungroup()

ggplot() +
  geom_ribbon(data = VBG, aes(cohort, k_median, ymin = k_lwr, ymax = k_upr,
                              fill = "All k's"), alpha = 0.4) +
  geom_ribbon(data = VBG_filter, aes(cohort, k_median, ymin = k_lwr, ymax = k_upr,
                                     fill = "Filtered"), alpha = 0.4) +
  geom_line(data = VBG, aes(cohort, k_median, color = "All k's")) + 
  geom_line(data = VBG_filter, aes(cohort, k_median, color = "Filtered")) + 
  guides(fill = FALSE) +
  facet_wrap(~area)

ggplot() +
  geom_line(data = VBG, aes(cohort, k_median, color = "All k's")) + 
  geom_line(data = VBG_filter, aes(cohort, k_median, color = "Filtered")) + 
  guides(fill = FALSE) +
  facet_wrap(~area)

# No difference at all really. We should probably just discuss that with this model, achieving biologically reasonable parameter values and a good fit to data are sometimes two different things. In our case, we just want a representative value of the growth (as in time to reach average maximum size in the population) that accounts for the entire growth trajectory of an individual, for each area and cohort.
```

## Add GAM-predicted temperature to growth models

```{r}
pred_temp <- read_csv(paste0(home, "/output/gam_predicted_temps.csv")) |> 
  rename(cohort = year)

VBG <- VBG |> left_join(pred_temp, by = c("area", "cohort"))

# save data for map-plot
cohort_sample_size <- IVBG |>
  group_by(area, cohort) |> 
  summarise(n = n()) # individuals, not samples!
  
VBG <- left_join(VBG, cohort_sample_size, by = c("cohort", "area"))

write_csv(VBG, paste0(home, "/output/vbg.csv"))
```

## Plot VBGE fits

```{r}
# Sample 50 IDs per area and plot their data and VBGE fits
set.seed(4)
ids <- IVBG |> distinct(ID, .keep_all = TRUE) |> group_by(area) |> slice_sample(n = 30)
#ids |> ungroup() |> group_by(area) |> summarise(n = length(unique(ID))) |> arrange(n)

IVBG2 <- IVBG |>
  filter(ID %in% ids$ID) |> 
  distinct(ID, .keep_all = TRUE) |> 
  select(ID, k, linf)
 
d2 <- d |>
  ungroup() |>
  filter(ID %in% ids$ID) |>
  left_join(IVBG2, by = "ID") |>
  mutate(length_mm_pred = linf*(1-exp(-k*age_bc)))
 
ggplot(d2, aes(age_bc, length_mm, group = ID, color = ID)) +
  geom_jitter(width = 0.3, height = 0, alpha = 0.5, size = 0.4) +
  geom_line(aes(age_bc, length_mm_pred, group = ID, color = ID),
            inherit.aes = FALSE, alpha = 0.8, linewidth = 0.3) + 
  guides(color = "none") +
  scale_color_viridis(discrete = TRUE, name = "Area", option = "cividis") +
  labs(x = "Age (yrs)", y = "Length (mm)") +
  facet_wrap(~factor(area, levels = (arrange(area_attr, desc(lat)))$area), ncol = 3)

ggsave(paste0(home, "/figures/vb_fits.pdf" ), width = 17, height = 17, units = "cm")

rugs <- IVBG |> 
  group_by(area) |> 
  summarise(median_k = median(k),
            median_linf = median(linf))

k <- IVBG |> 
  ggplot(aes(k, color = factor(area, order$area))) + 
  geom_density(alpha = 0.4, fill = NA, adjust = 1.5) +  
  scale_color_manual(values = colors, name = "Area") +
  coord_cartesian(expand = 0) + 
  labs(x = expression(italic(k))) +
  geom_rug(data = rugs, aes(median_k)) +
  theme(legend.position = c(0.9, 0.5))

linf <- IVBG |> 
  ggplot(aes(linf, color = factor(area, order$area))) + 
  geom_density(alpha = 0.4, fill = NA, adjust = 1.5) +  
  scale_color_manual(values = colors, name = "Area") +
  coord_cartesian(expand = 0, xlim = c(0, 2000)) + 
  labs(x = expression(paste(italic(L[infinity]), " [cm]"))) +
  geom_rug(data = rugs, aes(median_linf)) +
  guides(color = "none")

k / linf

ggsave(paste0(home, "/figures/vb_pars.pdf" ), width = 17, height = 17, units = "cm")
```

## Fit Sharpe-Schoolfield model to K

By area

```{r}
model <- 'sharpeschoolhigh_1981'

# Get starting values on full dataset for Sharpe-Schoolfield model
dat <- VBG |>
  select(k_median, temp, area) |>
  rename(rate = k_median)

lower <- get_lower_lims(dat$temp, dat$rate, model_name = model)
upper <- get_upper_lims(dat$temp, dat$rate, model_name = model)
start <- get_start_vals(dat$temp, dat$rate, model_name = model)
  
# Sharpe-Schoolfield model fit to data for each area
preds <- NULL
pred <- NULL

for(a in unique(dat$area)) {
  
  # Get data
  dd <- dat |> filter(area == a)
  
  # Fit model
  fit <- nls_multstart(
    rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 8),
    data = dd,
    iter = c(3, 3, 3, 3),
    start_lower = start*0.5,
    start_upper = start*2,
    lower = lower,
    upper = upper,
    supp_errors = 'Y'
    )
  
  # Make predictions on new data
  new_data <- data.frame(temp = seq(min(dd$temp), max(dd$temp), length.out = 100))
  
  pred <- augment(fit, newdata = new_data) |> mutate(area = a)
  
  # Add to general data frame
  preds <- data.frame(rbind(preds, pred))
  
}
```

All areas pooled

```{r}
# One sharpe schoolfield fitted to all data
fit_all <- nls_multstart(
    rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 8),
    data = dat,
    iter = c(3, 3, 3, 3),
    start_lower = start*0.5,
    start_upper = start*2,
    lower = lower,
    upper = upper,
    supp_errors = 'Y'
    )

# Make predictions on new data
new_data_all <- data.frame(temp = seq(min(dat$temp), max(dat$temp), length.out = 100))

pred_all <- augment(fit_all, newdata = new_data_all) |> 
  mutate(area = "all")

# Add t_opt
kb <- 8.62e-05
data.frame(par = names(coef(fit_all)), est = coef(fit_all)) |> 
  pivot_wider(names_from = par, values_from = est) |> 
  summarise(t_opt = (eh*th) / (eh + kb*th*log((eh/e)-1)))
```

## Plot Sharpe Schoolfield fits

```{r}
# Plot growth coefficients by year and area against mean SST
p1 <- preds |>
  ggplot(aes(temp, .fitted, color = factor(area, order$area))) + 
  geom_point(data = dat, aes(temp, rate, color = factor(area, order$area)), size = 0.6) +
  geom_line(linewidth = 1) +
  geom_line(data = pred_all, aes(temp, .fitted), linewidth = 1, inherit.aes = FALSE, linetype = 2) +
  scale_color_manual(values = colors, name = "Area") +
  guides(color = guide_legend(override.aes = list(size = 1))) +
  scale_x_continuous(breaks = seq(-5, 20, 1)) +
  labs(x = "Temperature (째C)",
       y = "von Bertalanffy growth coefficient (*k*)") +
  theme(axis.title.y = ggtext::element_markdown())

p1 + facet_wrap(~factor(area, levels = (arrange(area_attr, desc(lat)))$area))
  
p1

ggsave(paste0(home, "/figures/sharpe_school.pdf" ), width = 15, height = 15, units = "cm")
```

## Extra analysis

### Temperature in growing season instead of all year average
#### Fit Sharpe-Schoolfield model to K
By area

```{r}
model <- 'sharpeschoolhigh_1981'

# Get starting values on full dataset for Sharpe-Schoolfield model
dat2 <- VBG |>
  select(k_median, temp_gs, area) |>
  rename(rate = k_median,
         temp = temp_gs) # growing season! not annual average!

lower <- get_lower_lims(dat2$temp, dat2$rate, model_name = model)
upper <- get_upper_lims(dat2$temp, dat2$rate, model_name = model)
start <- get_start_vals(dat2$temp, dat2$rate, model_name = model)
  
# Sharpe-Schoolfield model fit to data for each area
preds2 <- NULL
pred2 <- NULL

for(a in unique(dat2$area)) {
  
  # Get data
  dd <- dat2 |> filter(area == a)
  
  # Fit model
  fit2 <- nls_multstart(
    rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 8),
    data = dd,
    iter = c(3, 3, 3, 3),
    start_lower = start*0.5,
    start_upper = start*2,
    lower = lower,
    upper = upper,
    supp_errors = 'Y'
    )
  
  # Make predictions on new data
  new_data2 <- data.frame(temp = seq(min(dd$temp), max(dd$temp), length.out = 100))
  
  pred2 <- augment(fit2, newdata = new_data2) |> mutate(area = a)
  
  # Add to general data frame
  preds2 <- data.frame(rbind(preds2, pred2))
  
}
```

All areas pooled

```{r}
# One sharpe schoolfield fitted to all data
fit_all2 <- nls_multstart(
    rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 8),
    data = dat2,
    iter = c(3, 3, 3, 3),
    start_lower = start*0.5,
    start_upper = start*2,
    lower = lower,
    upper = upper,
    supp_errors = 'Y'
    )

# Make predictions on new data
new_data_all2 <- data.frame(temp = seq(min(dat2$temp), max(dat2$temp), length.out = 100))

pred_all2 <- augment(fit_all2, newdata = new_data_all2) |> 
  mutate(area = "all")

# Add t_opt
kb <- 8.62e-05
data.frame(par = names(coef(fit_all2)), est = coef(fit_all2)) |> 
  pivot_wider(names_from = par, values_from = est) |> 
  summarise(t_opt = (eh*th) / (eh + kb*th*log((eh/e)-1)))
```

## Plot Sharpe Schoolfield fits

```{r}
# Plot growth coefficients by year and area against mean SST
preds2 |>
  ggplot(aes(temp, .fitted, color = factor(area, order$area))) + 
  geom_point(data = dat2, aes(temp, rate, color = factor(area, order$area)), size = 0.6) +
  geom_line(linewidth = 1) +
  geom_line(data = pred_all2, aes(temp, .fitted), linewidth = 1, inherit.aes = FALSE, linetype = 2) +
  scale_color_manual(values = colors, name = "Area") +
  guides(color = guide_legend(override.aes = list(size = 1))) +
  scale_x_continuous(breaks = seq(-5, 20, 1)) +
  labs(x = "Temperature (째C)",
       y = "von Bertalanffy growth coefficient (*k*)") +
  theme(axis.title.y = ggtext::element_markdown())
```

### Linf instead of k
By area

```{r}
model <- 'sharpeschoolhigh_1981'

# Get starting values on full dataset for Sharpe-Schoolfield model
dat3 <- VBG |>
  select(linf_median, temp, area) |>
  rename(rate = linf_median)

lower <- get_lower_lims(dat3$temp, dat3$rate, model_name = model)
upper <- get_upper_lims(dat3$temp, dat3$rate, model_name = model)
start <- get_start_vals(dat3$temp, dat3$rate, model_name = model)
  
# Sharpe-Schoolfield model fit to data for each area
preds3 <- NULL
pred3 <- NULL

for(a in unique(dat3$area)) {
  
  # Get data
  dd <- dat3 |> filter(area == a)
  
  # Fit model
  fit3 <- nls_multstart(
    rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 8),
    data = dd,
    iter = c(3, 3, 3, 3),
    start_lower = start*0.5,
    start_upper = start*2,
    lower = lower,
    upper = upper,
    supp_errors = 'Y'
    )
  
  # Make predictions on new data
  new_data3 <- data.frame(temp = seq(min(dd$temp), max(dd$temp), length.out = 100))
  
  pred3 <- augment(fit3, newdata = new_data3) |> mutate(area = a)
  
  # Add to general data frame
  preds3 <- data.frame(rbind(preds3, pred3))
  
}
```

All areas pooled

```{r}
# One sharpe schoolfield fitted to all data
fit_all3 <- nls_multstart(
    rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 8),
    data = dat3,
    iter = c(3, 3, 3, 3),
    start_lower = start*0.5,
    start_upper = start*2,
    lower = lower,
    upper = upper,
    supp_errors = 'Y'
    )

# Make predictions on new data
new_data_all3 <- data.frame(temp = seq(min(dat3$temp), max(dat3$temp), length.out = 100))

pred_all3 <- augment(fit_all3, newdata = new_data_all3) |> 
  mutate(area = "all")
```

## Plot Sharpe Schoolfield fits

```{r}
# Plot growth coefficients by year and area against mean SST
preds3 |>
  ggplot(aes(temp, .fitted, color = factor(area, order$area))) + 
  geom_point(data = dat3, aes(temp, rate, color = factor(area, order$area)), size = 0.6) +
  geom_line(linewidth = 1) +
  geom_line(data = pred_all3, aes(temp, .fitted), linewidth = 1, inherit.aes = FALSE, linetype = 2) +
  scale_color_manual(values = colors, name = "Area") +
  guides(color = guide_legend(override.aes = list(size = 1))) +
  scale_x_continuous(breaks = seq(-5, 20, 1)) +
  labs(x = "Temperature (째C)",
       y = expression(paste(italic(L[infinity]), " [cm]"))) +
  theme(axis.title.y = ggtext::element_markdown())
```

```{r}
ggplot(VBG, aes(linf_median, k_median)) + 
  geom_point()
```

