---
title: "Fit VBGE models to back-calculated length-at-age"
author: "Max Lindmark, Jan Ohlberger, Anna Gårdmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 80%
editor: source
execute: 
  echo: true
  eval: true
  cache: false
---

## Load libraries

```{r}
#| cache: false
library(tidyverse)
library(tidylog)
library(broom)
library(RColorBrewer)
library(viridis)
library(minpack.lm)
library(patchwork)
library(ggtext)
library(brms)
library(modelr)
library(tidybayes)
library(ggridges)

# devtools::install_github("seananderson/ggsidekick") ## not on CRAN 
library(ggsidekick); theme_set(theme_sleek())

# Load functions
home <- here::here()
fxn <- list.files(paste0(home, "/R/functions"))
invisible(sapply(FUN = source, paste0(home, "/R/functions/", fxn)))
```

Load cache

```{r}
# qwraps2::lazyload_cache_dir(path = paste0(home, "/R/analyze-data/02-fit-vbge_cache/html"))
```

## Read and trim data

```{r}
#| cache: false
d <- #read_csv(paste0(home, "/data/for-analysis/dat.csv")) %>% 
  read_csv("https://raw.githubusercontent.com/maxlindmark/perch-growth/master/data/for-analysis/dat.csv") %>% 
  filter(age_ring == "Y", # use only length-at-age by filtering on age_ring
         !area %in% c("VN", "TH")) 

# Minimum number of observations per area and cohort
d$area_cohort <- as.factor(paste(d$area, d$cohort))

d <- d %>%
  group_by(area_cohort) %>% 
  filter(n() > 100)

# Minimum number of observations per area, cohort, and age
d$area_cohort_age <- as.factor(paste(d$area, d$cohort, d$age_bc))

d <- d %>%
  group_by(area_cohort_age) %>% 
  filter(n() > 10)

# Minimum number of cohorts in a given area
cnt <- d %>%
  group_by(area) %>%
  summarise(n = n_distinct(cohort)) %>%
  filter(n >= 10)

d <- d[d$area %in% cnt$area, ]

# Plot cleaned data
ggplot(d, aes(age_bc, length_mm)) +
  geom_jitter(size = 0.1, height = 0, alpha = 0.1) +
  scale_x_continuous(breaks = seq(20)) +
  theme(axis.text.x = element_text(angle = 0)) +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 15)) +
  labs(x = "Age", y = "Length (mm)") +
  guides(color = "none") + 
  facet_wrap(~area, scale = "free_x")

# Longitude and latitude attributes for each area
area <- c("BS", "BT", "FB", "FM", "HO", "JM", "MU", "RA", "SI_EK", "SI_HA", "TH")
nareas <- length(area)
lat <- c(60, 60.4, 60.3, 60.5, 63.7, 58, 59, 65.9, 57.3, 57.4, 56.1)
lon <- c(21.5, 18.1, 19.5, 18, 20.9, 16.8, 18.1, 22.3, 16.6, 16.7, 15.9)
area_attr <- data.frame(cbind(area = area, lat = lat, lon = lon)) %>%
  mutate_at(c("lat", "lon"), as.numeric)

```

## Fit VBGE models

```{r}
# Get individual growth parameters (functions: VBGF/Gompertz,nls_out,fit_nls)
IVBG <- d %>%
  group_by(ID) %>%
  summarize(nls_out(fit_nls(length_mm, age_bc, min_nage = 5, model = "VBGF")))
```

```{r}
# TEST remove
t <- IVBG %>% drop_na()

t %>% 
  ggplot(aes(linf, k)) + 
  geom_point() + 
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")

t %>% 
  ggplot(aes(linf, 0.65*k*linf)) + 
  geom_point() + 
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")

broom::tidy(lm(0.65*k*linf ~ linf, data = t))
```


```{r}
#| cache: false
IVBG <- IVBG %>% drop_na(k) # The NAs are because the model didn't converge or because they were below the threshold age

IVBG <- IVBG %>%
  mutate(k_lwr = k - 1.96*k_se,
         k_upr = k + 1.96*k_se,
         linf_lwr = linf - 1.96*linf_se,
         linf_upr = linf + 1.96*linf_se,
         row_id = row_number())

# Plot all K's
IVBG %>%
  #filter(row_id < 5000) %>%
  ggplot(aes(row_id, k, ymin = k_lwr, ymax = k_upr)) +
  geom_point(alpha = 0.2) +
  geom_errorbar(alpha = 0.2) +
  NULL

# Plot all L_inf's
IVBG %>%
  #filter(row_id < 5000) %>%
  ggplot(aes(row_id, linf, ymin = linf_lwr, ymax = linf_upr)) +
  geom_point(alpha = 0.2) +
  geom_errorbar(alpha = 0.2) +
  NULL

# We can also consider removing individuals where the SE of k is larger than the fit
IVBG %>%
  #mutate(keep = ifelse(k > quantile(k_se, probs = 0.95), "N", "Y")) %>%
  mutate(keep = ifelse(k < k_se, "N", "Y")) %>%
  #filter(row_id < 10000) %>%
  ggplot(aes(row_id, k, ymin = k_lwr, ymax = k_upr, color = keep)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~keep, ncol = 1) +
  geom_errorbar(alpha = 0.5) +
  NULL

# Add back cohort and area variables
IVBG <- IVBG %>% 
  left_join(d %>% select(ID, area_cohort)) %>% 
  separate(area_cohort, into = c("area", "cohort"), remove = TRUE, sep = " ") %>% 
  mutate(cohort = as.numeric(cohort))

# Compare how the means and quantiles differ depending on this filtering
IVBG_filter <- IVBG %>%
  drop_na(k_se) %>%
  #filter(k_se < quantile(k_se, probs = 0.95)) %>% 
  filter(k_se < k)

# Summarize growth coefficients by cohort and area
VBG <- IVBG %>%
  filter(k_se < k) %>% # new!
  group_by(cohort, area) %>%
  summarize(k_mean = mean(k, na.rm = T),
            k_median = quantile(k, prob = 0.5, na.rm = T),
            linf_lwr = quantile(linf, prob = 0.05, na.rm = T),
            linf_median = quantile(linf, prob = 0.5, na.rm = T),
            linf_upr = quantile(linf, prob = 0.95, na.rm = T),
            k_lwr = quantile(k, prob = 0.05, na.rm = T),
            k_upr = quantile(k, prob = 0.95, na.rm = T)) %>% 
  ungroup()

VBG_filter <- IVBG_filter %>%
  group_by(cohort, area) %>%
  summarize(k_mean = mean(k, na.rm = T),
            k_median = quantile(k, prob = 0.5, na.rm = T),
            k_lwr = quantile(k, prob = 0.05, na.rm = T),
            k_upr = quantile(k, prob = 0.95, na.rm = T)) %>% 
  ungroup()

ggplot() +
  geom_ribbon(data = VBG, aes(cohort, k_median, ymin = k_lwr, ymax = k_upr,
                              fill = "All k's"), alpha = 0.4) +
  geom_ribbon(data = VBG_filter, aes(cohort, k_median, ymin = k_lwr, ymax = k_upr,
                                     fill = "Filtered"), alpha = 0.4) +
  geom_line(data = VBG, aes(cohort, k_median, color = "All k's")) + 
  geom_line(data = VBG_filter, aes(cohort, k_median, color = "Filtered"), linetype = 2) + 
  guides(fill = "none") +
  facet_wrap(~area)

ggplot() +
  geom_line(data = VBG, aes(cohort, k_median, color = "All k's")) + 
  geom_line(data = VBG_filter, aes(cohort, k_median, color = "Filtered"), linetype = 2) + 
  guides(fill = "none") +
  facet_wrap(~area)

# No difference at all really. We should probably just discuss that with this model, achieving biologically reasonable parameter values and a good fit to data are sometimes two different things. In our case, we just want a representative value of the growth (as in time to reach average maximum size in the population) that accounts for the entire growth trajectory of an individual, for each area and cohort.
```

Calculate sample size

```{r}
sample_size <- IVBG %>%
  group_by(area) %>% 
  summarise(n_cohort = length(unique(cohort)),
            min_cohort = min(cohort),
            max_cohort = max(cohort),
            n_individuals = length(unique(ID)),
            n_data_points = n())

sample_size

sample_size %>%
  ungroup() %>%
  summarise(sum_ind = sum(n_individuals), sum_n = sum(n_data_points))

write_csv(sample_size, paste0(home, "/output/sample_size.csv"))
```

## Add GAM-predicted temperature to growth models

```{r}
#| cache: false
pred_temp <- read_csv(paste0(home, "/output/gam_predicted_temps.csv")) %>% 
  rename(cohort = year)

VBG <- VBG %>%
  left_join(pred_temp, by = c("area", "cohort"))

# Save data for map-plot
cohort_sample_size <- IVBG %>%
  group_by(area, cohort) %>% 
  summarise(n = n()) # individuals, not samples!
  
VBG <- left_join(VBG, cohort_sample_size, by = c("cohort", "area"))

write_csv(VBG, paste0(home, "/output/vbg.csv"))

# Calculate the plotting order (also used for map plot)
order <- VBG %>%
  ungroup() %>%
  group_by(area) %>%
  summarise(min_temp = min(temp)) %>%
  arrange(desc(min_temp))

order

write_csv(order, paste0(home, "/output/ranked_temps.csv"))

nareas <- length(unique(order$area)) + 2 # to skip the brightest colors that are hard to read
colors <- colorRampPalette(brewer.pal(name = "RdYlBu", n = 10))(nareas)[-c(6,7)]
```

## Plot VBGE fits

```{r}
#| cache: false
# Sample 30 IDs per area and plot their data and VBGE fits
set.seed(4)
ids <- IVBG %>% distinct(ID, .keep_all = TRUE) %>% group_by(area) %>% slice_sample(n = 30)

IVBG2 <- IVBG %>%
  filter(ID %in% ids$ID) %>% 
  distinct(ID, .keep_all = TRUE) %>% 
  select(ID, k, linf)
 
d2 <- d %>%
  ungroup() %>%
  filter(ID %in% ids$ID) %>%
  left_join(IVBG2, by = "ID") %>%
  mutate(length_mm_pred = linf*(1-exp(-k*age_bc)))
 
fits_ind <- ggplot(d2, aes(age_bc, length_mm, group = ID, color = ID)) +
  geom_jitter(width = 0.3, height = 0, alpha = 0.5, size = 0.4) +
  geom_line(aes(age_bc, length_mm_pred, group = ID, color = ID),
            inherit.aes = FALSE, alpha = 0.8, linewidth = 0.3) + 
  guides(color = "none") +
  scale_color_viridis(discrete = TRUE, name = "Site", option = "cividis") +
  labs(x = "Age (years)", y = "Length (mm)") +
  facet_wrap(~factor(area, levels = (arrange(area_attr, desc(lat)))$area), ncol = 5) + 
  NULL

k <- IVBG %>% 
  ggplot(aes(factor(area, order$area), k, fill = factor(area, order$area))) + 
  coord_cartesian(ylim = c(0, 0.7)) +
  geom_violin(alpha = 0.8, color = NA) +  
  geom_boxplot(outlier.color = NA, width = 0.2, alpha = 0.9, fill = NA, size = 0.4) +
  scale_fill_manual(values = colors, name = "Site") +
  guides(fill = "none", color = "none") +
  labs(x = "Site", y = expression(italic(k))) + 
  coord_flip()

linf <- IVBG %>% 
  filter(linf < 2300) %>% 
  filter(linf > 130) %>% 
  ggplot(aes(factor(area, order$area), linf, fill = factor(area, order$area))) + 
  geom_violin(alpha = 0.8, color = NA) +  
  geom_boxplot(outlier.color = NA, width = 0.1, alpha = 0.9, fill = NA, size = 0.4) +
  coord_cartesian(ylim = c(0, 2000)) +
  scale_fill_manual(values = colors, name = "Site") +
  guides(fill = "none", color = "none") +
  labs(x = "", y = expression(paste(italic(L[infinity]), " [mm]"))) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  coord_flip()

#fits_ind / (k + linf) + plot_annotation(tag_levels = "A") + plot_layout(heights = c(1, 1.8))

k2 <- k + coord_cartesian(ylim = c(0, 0.65))
fits_ind / k2 + plot_annotation(tag_levels = "A") #+ plot_layout(heights = c(1, 1.8))

ggsave(paste0(home, "/figures/vb_pars_fits.pdf" ), width = 17, height = 18, units = "cm")
```

## Fit Sharpe-Schoolfield model to k

```{r}
dat <- VBG %>%
  select(k_median, temp, area) %>%
  rename(rate = k_median)
```

```{r}
#| cache: false
# TEST REMOVE
# Testing different version of growth metrics!
dat <- VBG %>%
  select(k_median, temp, area) %>%
  rename(rate = k_median)
  # select(k_median, linf_median, temp, area) %>%
  # mutate(rate = k_median*linf_median*0.65)

model <- 'sharpeschoolhigh_1981'

# Get starting values on full dataset for Sharpe-Schoolfield model
# dat <- VBG %>%
#   select(k_median, temp, area) %>%
#   rename(rate = k_median)

lower <- get_lower_lims(dat$temp, dat$rate, model_name = model)
upper <- get_upper_lims(dat$temp, dat$rate, model_name = model)
start <- get_start_vals(dat$temp, dat$rate, model_name = model)
  
# Sharpe-Schoolfield model fit to data for each area
preds <- NULL
pred <- NULL
pars <- list()
t_opts <- list()

for(a in unique(dat$area)) {
  
  # Get data
  dd <- dat %>% filter(area == a)
  
  # Fit model
  fit <- nls_multstart(
    rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 8),
    data = dd,
    iter = c(3, 3, 3, 3),
    start_lower = start*0.5,
    start_upper = start*2,
    lower = lower,
    upper = upper,
    supp_errors = 'Y'
    )
  
  # Make predictions on new data
  new_data <- data.frame(temp = seq(min(dd$temp), max(dd$temp), length.out = 100))
  
  pred <- augment(fit, newdata = new_data) %>% mutate(area = a)
  
  # Add to general data frame
  preds <- data.frame(rbind(preds, pred))
  
  # Add parameter estimates
  pars[[a]] <- summary(fit)$coefficients %>%
    as.data.frame() %>%
    rownames_to_column(var = "parameter") %>%
    mutate(area = a)
  
  # Extract t_topt
  t_opts[[a]] <- data.frame(t_opt = get_topt(fit),
                            area = a)

}

# Plot
p1 <- preds %>%
  ggplot(aes(temp, .fitted, color = factor(area, order$area))) + 
  geom_point(data = dat, aes(temp, rate, color = factor(area, order$area)), size = 1, alpha = 0.8) +
  geom_line(linewidth = 1) +
  #geom_line(data = pred_all, aes(temp, .fitted), linewidth = 1, inherit.aes = FALSE, linetype = 2) +
  scale_color_manual(values = colors, name = "Site") +
  guides(color = guide_legend(nrow = 1, reverse = TRUE, title.position = "top", title.hjust = 0.5,
                              override.aes = list(size = 1))) +
  scale_x_continuous(breaks = seq(-5, 20, 2)) +
  labs(x = "Temperature (°C)",
       y = "von Bertalanffy growth coefficient (*k*)") +
  theme(axis.title.y = ggtext::element_markdown(),
        legend.position = "bottom",
        legend.direction = "horizontal")




# Now do the A
# Testing different version of growth metrics!
# Compare how the means and quantiles differ depending on this filtering
IVBG_filter <- IVBG %>%
  drop_na(k_se) %>%
  drop_na(linf) %>%
  filter(k_se < 0.5*k) %>% 
  filter(linf_se < 0.5*linf)

# Summarize growth coefficients by cohort and area
VBG <- IVBG_filter %>%
  filter(k_se < k) %>% # new!
  group_by(cohort, area) %>%
  summarize(k_mean = mean(k, na.rm = T),
            k_median = quantile(k, prob = 0.5, na.rm = T),
            linf_lwr = quantile(linf, prob = 0.05, na.rm = T),
            linf_median = quantile(linf, prob = 0.5, na.rm = T),
            linf_upr = quantile(linf, prob = 0.95, na.rm = T),
            k_lwr = quantile(k, prob = 0.05, na.rm = T),
            k_upr = quantile(k, prob = 0.95, na.rm = T)) %>% 
  ungroup()

VBG %>% 
  mutate(iteration = seq_len(n())) %>% 
  filter(linf_median < 600) %>% 
  ggplot(aes(iteration, linf_median, ymax = linf_upr, ymin = linf_lwr)) + 
  geom_pointrange()
  
VBG <- VBG %>% 
  filter(linf_upr < 2*linf_median)

VBG <- VBG %>%
  left_join(pred_temp, by = c("area", "cohort"))

dat2 <- VBG %>%
  # select(k_median, temp, area) %>%
  # rename(rate = k_median)
  select(k_median, linf_median, temp, area) %>%
  mutate(rate = k_median*linf_median*0.65)

lower <- get_lower_lims(dat2$temp, dat2$rate, model_name = model)
upper <- get_upper_lims(dat2$temp, dat2$rate, model_name = model)
start <- get_start_vals(dat2$temp, dat2$rate, model_name = model)
  
# Sharpe-Schoolfield model fit to data for each area
preds2 <- NULL
preds2_long <- NULL
pred2 <- NULL
pred2_long <- NULL
t_opts2 <- list()
t_opts2_long <- list()

for(a in unique(dat2$area)) {
  
  # Get data
  dd <- dat2 %>% filter(area == a)
  
  # Fit model
  fit <- nls_multstart(
    rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 8),
    data = dd,
    iter = c(3, 3, 3, 3),
    start_lower = start*0.5,
    start_upper = start*2,
    lower = lower,
    upper = upper,
    supp_errors = 'Y'
    )
  
  # Make predictions on new data
  new_data <- data.frame(temp = seq(min(dd$temp), max(dd$temp), length.out = 100))
  new_data_long <- data.frame(temp = seq(5, 35), length.out = 100)
  
  pred2 <- augment(fit, newdata = new_data) %>% mutate(area = a)
  pred2_long <- augment(fit, newdata = new_data_long) %>% mutate(area = a)
  
  # Add to general data frame
  preds2 <- data.frame(rbind(preds2, pred2))
  preds2_long <- data.frame(rbind(preds2_long, pred2_long))
    
  # Extract t_topt
  t_opts2[[a]] <- data.frame(t_opt = get_topt(fit),
                             area = a)
  
  t_opts2_long[[a]] <- data.frame(t_opt_long = get_topt(fit),
                                  area = a)

}

# Plot
p2 <- preds2 %>%
  ggplot(aes(temp, .fitted, color = factor(area, order$area))) + 
  geom_point(data = dat2, aes(temp, rate, color = factor(area, order$area)), size = 1, alpha = 0.8) +
  geom_line(linewidth = 1) +
  #geom_line(data = pred_all, aes(temp, .fitted), linewidth = 1, inherit.aes = FALSE, linetype = 2) +
  scale_color_manual(values = colors, name = "Site") +
  guides(color = guide_legend(nrow = 1, reverse = TRUE, title.position = "top", title.hjust = 0.5,
                              override.aes = list(size = 1))) +
  scale_x_continuous(breaks = seq(-5, 20, 2)) +
  labs(x = "Temperature (°C)",
       y = "A") +
  theme(axis.title.y = ggtext::element_markdown(),
        legend.position = "bottom",
        legend.direction = "horizontal")# + 

  #facet_wrap(~area)

p1 / p2

t_opts2_df <- bind_rows(t_opts2)
t_opts2_long_df <- bind_rows(t_opts2_long_df)

pred_temp2 <- pred_temp %>% summarise(mean_temp = mean(temp), .by = area)

t_opts2_long_df <- preds2_long %>%
  group_by(area) %>% 
  filter(.fitted == max(.fitted)) %>% 
  ungroup() %>% 
  left_join(pred_temp2)

library(ggrepel)
ggplot(t_opts2_df, aes(mean_temp, t_opt)) + 
  geom_abline(color = "tomato3") + 
  #geom_smooth(method = "lm", color = "steelblue") + 
  #theme(aspect.ratio = 1) + 
  #geom_point() +
  geom_point(data = t_opts2_long_df, aes(mean_temp, temp)) + 
  geom_smooth(data = t_opts2_long_df, aes(mean_temp, temp),
              method = "lm", color = "steelblue", se = FALSE) + 
  #geom_label_repel(aes(label = area)) +
  ylim(6, 16) + 
  xlim(6, 16)
```


## Can we fit a single Sharpe Schoolfield with area-specific parameters with `brms`?

```{r}
# Again, here are the data we are fitting:
ggplot(dat, aes(temp, rate, color = factor(area, order$area))) + 
  geom_point(data = dat, aes(temp, rate, color = factor(area, order$area)), size = 0.6)  

hist(dat$rate)

# Here's the equation (from the r package rTPC):
# > sharpeschoolhigh_1981
# function (temp, r_tref, e, eh, th, tref) 
# {
#     tref <- 273.15 + tref
#     k <- 8.62e-05
#     boltzmann.term <- r_tref * exp(e/k * (1/tref - 1/(temp + 
#         273.15)))
#     inactivation.term <- 1/(1 + exp(eh/k * (1/(th + 273.15) - 
#         1/(temp + 273.15))))
#     return(boltzmann.term * inactivation.term)
# }

# Add in fixed parameters
dat$bk <- 8.62e-05
dat$tref <- 8 + 273.15

# (better visualization including bounds further down)
n = 10000
hist(rnorm(0.8, 1, n = n), main = "e", xlim = c(0, 5))
hist(rnorm(2, 1, n = n), main = "eh", xlim = c(0, 7))
hist(rnorm(0.25, 0.5, n = n), main = "r_tref", xlim = c(0, 2.5))
hist(rnorm(12, 5, n = n), main = "th", xlim = c(5, 30))

# These work nicely with the pp_check
prior <- c(prior(normal(0.8, 1), nlpar = "e", lb = 0),
           prior(normal(2, 1), nlpar = "eh", lb = 0),
           prior(normal(0.3, 0.5), nlpar = "rtref", lb = 0),
           prior(normal(10, 5), nlpar = "th", lb = 0))

# Now to a prior predictive check
fit_prior <- brm(
  bf(rate ~ (rtref * exp(e/bk * (1/tref - 1/(temp + 273.15)))) / (1 + exp(eh/bk * (1/(th + 273.15) - 1/(temp + 273.15)))),
     rtref + e + eh + th ~ 1, 
     nl = TRUE),
  data = dat,
  cores = 2,
  chains = 2,
  iter = 1500,
  sample_prior = "only",
  seed = 9,
  prior = prior
)

# Global prior predictive check in relation to data. Doesn't loo too informative...
dat %>%
  data_grid(temp = seq_range(temp, n = 51)) %>% 
  ungroup() %>% 
  mutate(bk = 8.62e-05,
         tref = 8 + 273.15) %>% 
  add_epred_draws(fit_prior) %>% 
  ggplot(aes(temp, y = .epred)) +
  stat_lineribbon(.width = c(0.75), alpha = 0.3, color = "black", fill = "black") + 
  geom_point(data = dat, aes(temp, rate, color = factor(area, order$area)), size = 0.6) +
  scale_color_manual(values = colors, name = "Site") +
  labs(y = "Expectation of the posterior predictive distribution", x = "Temperature (°C)") +
  NULL

ggsave(paste0(home, "/figures/supp/prior_predictive_check.pdf" ), width = 15, height = 11, units = "cm")

# Now make sure it converges with real data but no random effects
fit_global <- brm(
  bf(rate ~ (rtref * exp(e/bk * (1/tref - 1/(temp + 273.15)))) / (1 + exp(eh/bk * (1/(th + 273.15) - 1/(temp + 273.15)))),
     rtref + e + eh + th ~ 1,
     nl = TRUE),
  data = dat,
  cores = 4,
  chains = 4,
  iter = 4000, 
  seed = 9,
  sample_prior = "yes",
  family = student,
  prior = prior
  )

plot(fit_global)
pp_check(fit_global)
plot(conditional_effects(fit_global, effect = "temp"), points = TRUE)

# Plot prior vs posterior
post <- fit_global %>%
  as_draws_df() %>%
  dplyr::select(b_rtref_Intercept, b_e_Intercept, b_eh_Intercept, b_th_Intercept) %>% 
  rename(r_tref = b_rtref_Intercept, 
         e = b_e_Intercept,
         eh = b_eh_Intercept,
         th = b_th_Intercept) %>% 
  pivot_longer(everything(), names_to = "parameter") %>% 
  mutate(type = "Posterior")

prior_draws <- fit_global %>%
  as_draws_df() %>%
  dplyr::select(prior_b_rtref, prior_b_e, prior_b_eh, prior_b_th) %>% 
  rename(r_tref = prior_b_rtref, 
         e = prior_b_e,
         eh = prior_b_eh,
         th = prior_b_th) %>% 
  pivot_longer(everything(), names_to = "parameter") %>% 
  mutate(type = "Prior")

dist <- bind_rows(prior_draws, post)

ggplot(dist, aes(value, fill = type)) +
  geom_density(color = NA, alpha = 0.5) +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~parameter, scales = "free") +
  theme(legend.position = c(0.35, 0.9)) +
  labs(fill = "")
```

Ok, seems to work with priors and everything. They are roughly as broad as can be and still have a fit. Now fit the full model, with random area effects, more iterations and chains!

```{r}
# Now look at random area effects!
nrow(dat)

# Student-t model with random effects
fitbs <- brm(
  bf(rate ~ (rtref * exp(e/bk * (1/tref - 1/(temp + 273.15)))) / (1 + exp(eh/bk * (1/(th + 273.15) - 1/(temp + 273.15)))),
     rtref + e + eh + th ~ 1 + (1|area),
     nl = TRUE),
  data = dat,
  family = student,
  cores = 4,
  chains = 4,
  iter = 5000,
  sample_prior = "yes",
  save_pars = save_pars(all = TRUE),
  seed = 9,
  prior = prior,
  control = list(adapt_delta = 0.99, max_treedepth = 12)
)

print(fitbs, digits = 4)
```

Inspect the model

```{r}
#| cache: false
# Check fit

library(performance)
fitbs
r2_bayes(fitbs)
r2_bayes(fit_global)

# qq-plots
p1 <- dat %>%
  add_predicted_draws(fitbs) %>%
  summarise(
    p_residual = mean(.prediction < rate),
    z_residual = qnorm(p_residual),
    .groups = "drop_last"
  ) %>%
  ggplot(aes(sample = z_residual)) +
  geom_qq() +
  geom_abline() + 
  labs(x = "Theoretical", y = "Sample") + 
  ggtitle("Mixed model") +
  theme(aspect.ratio = 1)

p2 <- dat %>%
  add_predicted_draws(fit_global) %>%
  summarise(
    p_residual = mean(.prediction < rate),
    z_residual = qnorm(p_residual),
    .groups = "drop_last"
  ) %>%
  ggplot(aes(sample = z_residual)) +
  geom_qq() +
  geom_abline() + 
  labs(x = "Theoretical", y = "Sample") + 
  ggtitle("Global model") +
  theme(aspect.ratio = 1)

p3 <- pp_check(fitbs) + theme(legend.position = c(0.1, 0.9)) + ggtitle("Mixed model")
p4 <- pp_check(fit_global) + theme(legend.position = c(0.1, 0.9)) + ggtitle("Global model")

(p1 + p2) / (p3 + p4)

ggsave(paste0(home, "/figures/supp/sharpe_brms_ppcheck_qq.pdf" ), width = 17, height = 17, units = "cm")

# Plot prior vs posterior
post <- fitbs %>%
  as_draws_df() %>%
  dplyr::select(b_rtref_Intercept, b_e_Intercept, b_eh_Intercept, b_th_Intercept) %>% 
  rename(r_tref = b_rtref_Intercept, 
         e = b_e_Intercept,
         eh = b_eh_Intercept,
         th = b_th_Intercept) %>% 
  pivot_longer(everything(), names_to = "parameter") %>% 
  mutate(type = "Posterior")

prior_draws <- fitbs %>%
  as_draws_df() %>%
  dplyr::select(prior_b_rtref, prior_b_e, prior_b_eh, prior_b_th) %>% 
    rename(r_tref = prior_b_rtref, 
           e = prior_b_e,
           eh = prior_b_eh,
           th = prior_b_th) %>% 
  pivot_longer(everything(), names_to = "parameter") %>% 
  mutate(type = "Prior")

dist <- bind_rows(prior_draws, post)

ggplot(dist, aes(value, fill = type)) +
  geom_density(color = NA, alpha = 0.5) +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~parameter, scales = "free") + 
  theme(legend.position = c(0.35, 0.9)) +
  labs(fill = "", x = "Value", y = "Density")

ggsave(paste0(home, "/figures/supp/sharpe_brms_prior.pdf" ), width = 17, height = 17, units = "cm")
```

Sensitivity with respect to priors

```{r}
# Set a bunch of priors with different sigmas
  # prior1 <- c(prior(normal(0.8, 1), nlpar = "e", lb = 0),
  #             prior(normal(2, 1), nlpar = "eh", lb = 0),
  #             prior(normal(0.3, 0.5), nlpar = "rtref", lb = 0),
  #             prior(normal(10, 5), nlpar = "th", lb = 0))

# For each object, we increase sd by a factor 1.5 from previous
  prior2 <- c(prior(normal(0.8, 1.5), nlpar = "e", lb = 0),
              prior(normal(2, 1.5), nlpar = "eh", lb = 0),
              prior(normal(0.3, 0.75), nlpar = "rtref", lb = 0),
              prior(normal(10, 7.5), nlpar = "th", lb = 0))
  
  prior3 <- c(prior(normal(0.8, 3), nlpar = "e", lb = 0),
              prior(normal(2, 3), nlpar = "eh", lb = 0),
              prior(normal(0.3, 1.125), nlpar = "rtref", lb = 0),
              prior(normal(10, 11.25), nlpar = "th", lb = 0))
  
  prior4 <- c(prior(normal(0.8, 4.5), nlpar = "e", lb = 0),
              prior(normal(2, 4.5), nlpar = "eh", lb = 0),
              prior(normal(0.3, 1.6875), nlpar = "rtref", lb = 0),
              prior(normal(10, 16.875), nlpar = "th", lb = 0))
  
  prior5 <- c(prior(normal(0.8, 6.75), nlpar = "e", lb = 0),
              prior(normal(2, 6.75), nlpar = "eh", lb = 0),
              prior(normal(0.3, 2.53125), nlpar = "rtref", lb = 0),
              prior(normal(10, 25.3125), nlpar = "th", lb = 0))
  
  prior6 <- c(prior(normal(0.8, 10.125), nlpar = "e", lb = 0),
              prior(normal(2, 10.125), nlpar = "eh", lb = 0),
              prior(normal(0.3, 3.796875), nlpar = "rtref", lb = 0),
              prior(normal(10, 37.96875), nlpar = "th", lb = 0))

  priors <- list(prior2, prior3, prior4, prior5, prior6)

# List for storing posteriors and priors
dist_list <- list()
  
for(i in 1:5) {
  
  prior <- priors[[i]]
  
  fit_sens <- brm(
    bf(rate ~ (rtref * exp(e/bk * (1/tref - 1/(temp + 273.15)))) / (1 + exp(eh/bk * (1/(th + 273.15) - 1/(temp + 273.15)))),
       rtref + e + eh + th ~ 1,
       nl = TRUE),
    data = dat,
    cores = 4,
    chains = 4,
    iter = 4000, 
    seed = 9,
    sample_prior = "yes",
    family = student,
    prior = prior
    )
  
  # Plot prior vs posterior
  post <- fit_sens %>%
    as_draws_df() %>%
    dplyr::select(b_rtref_Intercept, b_e_Intercept, b_eh_Intercept, b_th_Intercept) %>% 
    rename(r_tref = b_rtref_Intercept, 
           e = b_e_Intercept,
           eh = b_eh_Intercept,
           th = b_th_Intercept) %>% 
    pivot_longer(everything(), names_to = "parameter") %>% 
    mutate(type = "Posterior")

  prior_draws <- fit_sens %>%
    as_draws_df() %>%
    dplyr::select(prior_b_rtref, prior_b_e, prior_b_eh, prior_b_th) %>% 
    rename(r_tref = prior_b_rtref, 
           e = prior_b_e,
           eh = prior_b_eh,
           th = prior_b_th) %>% 
    pivot_longer(everything(), names_to = "parameter") %>% 
    mutate(type = "Prior")
  
  dist <- bind_rows(prior_draws, post) %>% 
    mutate(iter = i)
  
  dist_list[[i]] <- dist
  
}

dist_df <- bind_rows(dist_list)

write_csv(dist_df, paste0(home, "/output/prior_sens.csv"))

ggplot(dist_df, aes(value, fill = type)) +
  geom_density(color = NA, alpha = 0.5) +
  scale_fill_brewer(palette = "Set1") +
  ggh4x::facet_grid2(iter~parameter, scales = "free", independent = "y") +
  #coord_cartesian(xlim = c(0, 50)) +
  labs(fill = "") + 
  labs(x = "Value", y = "Density")

ggsave(paste0(home, "/figures/supp/prior_sensi.pdf"), width = 17, height = 17, units = "cm")

ggplot(dist_df %>% filter(parameter == "r_tref"), aes(value, color = as.factor(iter), 
                    linetype = as.factor(type))) +
  geom_density(fill = NA) +
  coord_cartesian(xlim = c(0, 4)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~parameter, scales = "free") +
  labs(fill = "", linetype = "", color = "iteration")

ggplot(dist_df %>% filter(parameter == "e"), aes(value, color = as.factor(iter), 
                    linetype = as.factor(type))) +
  geom_density(fill = NA) +
  coord_cartesian(xlim = c(0, 12)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~parameter, scales = "free") +
  labs(fill = "", linetype = "", color = "iteration")

ggplot(dist_df %>% filter(parameter == "eh"), aes(value, color = as.factor(iter), 
                    linetype = as.factor(type))) +
  geom_density(fill = NA) +
  coord_cartesian(xlim = c(0, 10)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~parameter, scales = "free") +
  labs(fill = "", linetype = "", color = "iteration")

ggplot(dist_df %>% filter(parameter == "th"), aes(value, color = as.factor(iter), 
                    linetype = as.factor(type))) +
  geom_density(fill = NA) +
  coord_cartesian(xlim = c(0, 35)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~parameter, scales = "free") +
  labs(fill = "", linetype = "", color = "iteration")
  
dist_df %>% 
  group_by(iter, parameter, type) %>% 
  summarise(mean = mean(value),
            sd = sd(value)) %>% 
  ggplot(aes(iter, mean, color = type)) +
  geom_point(position = position_dodge(width = 0.2)) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0,
                position = position_dodge(width = 0.2)) +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~parameter, scales = "free")

dist_df %>% 
  group_by(iter, parameter, type) %>% 
  summarise(mean = mean(value)) %>% 
  ggplot(aes(iter, mean, color = type)) +
  geom_point(position = position_dodge(width = 0.2)) +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~parameter, scales = "free")

dist_df %>% 
  group_by(iter, parameter, type) %>% 
  summarise(mean = mean(value),
            sd = sd(value)) %>% 
  arrange(parameter) %>% 
  as.data.frame()
```

Calculate T_opt

```{r}
ts <- dat %>% 
  data_grid(temp = seq_range(temp, n = 100)) %>% 
  mutate(bk = 8.62e-05,
         tref = 8 + 273.15) %>% 
  add_epred_draws(fit_global, re_formula = NA, ndraws = 8000) 

p1 <- ggplot(ts) + 
  geom_line(aes(temp, y = .epred, group = .draw), alpha = .1) + 
  coord_cartesian(ylim = c(0.1, 0.4))

p2 <- ggplot(ts) + 
  geom_line(aes(temp, y = .epred, group = .draw), alpha = .1) +
  coord_cartesian(ylim = c(0.1, 0.4))

p1 + p2

# Compute quantiles across spaghetties
t_opt <- ts %>% 
  group_by(.draw) %>% 
  filter(.epred == max(.epred)) %>% 
  ungroup() %>% 
  filter(!temp == min(temp)) %>% # remove values where there was no optimum
  filter(!temp == max(temp))

# These are used in the fit-temp script
quantile(t_opt$temp, probs = c(0.025, 0.5, 0.975))

ggplot(t_opt, aes(temp)) + 
  geom_histogram(bins = length(unique(t_opt$temp))) + 
  theme_void()

# Make the main plot (conditional effect of temperature, with and without random effects)
# Predictions without random effects
glob <- dat %>%
  data_grid(temp = seq_range(temp, n = 100)) %>% 
  mutate(bk = 8.62e-05,
         tref = 8 + 273.15) %>% 
  add_epred_draws(fit_global, re_formula = NA) %>% #
  ungroup()

dat %>%
  group_by(area) %>%
  data_grid(temp = seq_range(temp, n = 100)) %>% 
  ungroup() %>% 
  mutate(bk = 8.62e-05,
         tref = 8 + 273.15) %>% 
  add_epred_draws(fitbs) %>% 
  ungroup() %>% 
  ggplot(aes(temp, y = .epred, color = factor(area, order$area))) +
  stat_lineribbon(data = glob, aes(temp, .epred), .width = c(0.9), inherit.aes = FALSE,
                 fill = "black", color = "black", alpha = 0.1) +
  geom_point(data = dat, aes(temp, rate, color = factor(area, order$area)), size = 1, alpha = 0.8) +
  stat_lineribbon(.width = c(0)) +
  stat_lineribbon(data = glob, aes(temp, .epred), .width = c(0), inherit.aes = FALSE,
                 color = "black", alpha = 0.9, linetype = 2) +
  coord_cartesian(expand = 0) +
  scale_color_manual(values = colors, name = "Site") +
  guides(fill = "none",
         color = guide_legend(nrow = 1, reverse = TRUE, title.position = "top", title.hjust = 0.5,
                              override.aes = list(size = 1))) +
  theme(axis.title.y = ggtext::element_markdown(),
        legend.position = "bottom",
        legend.direction = "horizontal") +
  labs(x = "Temperature (°C)",
       y = "von Bertalanffy growth coefficient (*k*)") + 
  NULL

ggsave(paste0(home, "/figures/sharpe_school_bayes.pdf"), width = 17, height = 17, units = "cm")
#ggsave(paste0(home, "/figures/for-talks/sharpe_school_bayes.pdf"), width = 14, height = 14, units = "cm")
```

### Area-specific T_opts as a ridgeplot

```{r}
#| cache: false
# Calculate T_opt by area
tsa <- dat %>% 
  data_grid(area = unique(dat$area),
            temp = seq_range(temp, n = 100)) %>% 
  mutate(bk = 8.62e-05,
         tref = 8 + 273.15) %>% 
  add_epred_draws(fitbs, re_formula = NULL, ndraws = 10000) 

t_opt_a <- tsa %>% 
  group_by(.draw, area) %>% 
  filter(.epred == max(.epred)) %>% 
  ungroup() %>% 
  filter(!temp == min(temp)) %>% # remove values where there was no optimum
  filter(!temp == max(temp))

t_opt_a_sum <- t_opt_a %>% 
  group_by(area) %>% 
  summarise(median = median(temp),
            lwr = quantile(temp, probs = 0.1),
            upr = quantile(temp, probs = 0.9))

# How many datapoints are below or above optimum by area and globally?
left_join(dat, t_opt_a_sum, by = "area") %>% 
  filter(temp < median) %>% 
  nrow()

left_join(dat, t_opt_a_sum, by = "area") %>% 
  mutate(above = ifelse(temp > median, "Y", "N")) %>% 
  group_by(area) %>% 
  count(above) %>% 
  pivot_wider(names_from = above, values_from = n) %>% 
  mutate(prop = Y/(Y+N)) %>% 
  arrange(desc(prop))

# Without impacted?
left_join(dat, t_opt_a_sum, by = "area") %>% 
  filter(!area %in% c("BT", "SI_EK")) %>% 
  filter(temp < median) %>% 
  nrow()

dat %>% 
  group_by(area) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))

# nrow(filter(dat, temp < 9.468607))
# nrow(dat)

ggplot(t_opt, aes(temp)) + 
  geom_histogram(bins = length(unique(t_opt$temp)))

# Compute quantiles across spaghetties
rect_dat <- data.frame(xmin = c(quantile(t_opt$temp, probs = c(0.05)), quantile(t_opt$temp, probs = c(0.1))),
                       xmax = c(quantile(t_opt$temp, probs = c(0.95)), quantile(t_opt$temp, probs = c(0.9))),
                       ymin = c(-Inf, -Inf),
                       ymax = c(Inf, Inf),
                       interval = c("90%", "80%"))

ridge <- ggplot(t_opt_a, aes(temp, factor(area, levels = rev(order$area)), fill = factor(area, levels = rev(order$area)))) + 
  geom_vline(xintercept = quantile(t_opt$temp, probs = c(0.5)), alpha = 0.6, linetype = 2, linewidth = 0.75) +
  geom_rect(data = rect_dat %>% filter(interval == "90%"), aes(xmin = xmin, xmax = xmax), ymin = -Inf, ymax = Inf, 
            inherit.aes = FALSE, fill = "grey", alpha = 0.35) +
  geom_rect(data = rect_dat %>% filter(interval == "80%"), aes(xmin = xmin, xmax = xmax), ymin = -Inf, ymax = Inf, 
            inherit.aes = FALSE, fill = "grey", alpha = 1) +
  geom_density_ridges(alpha = 0.7, color = NA, scale = 1,
                      stat = "binline", bins = 50) +
  scale_fill_manual(values = rev(colors), name = "Site") +
  scale_color_manual(values = rev(colors), name = "Site") +
  labs(x = "Temperature (°C)", y = "") +
  guides(fill = "none") +
  geom_errorbar(data = t_opt_a_sum, aes(xmin = lwr, xmax = upr, y = area, color = factor(area, levels = rev(order$area))), linewidth = 1,
                inherit.aes = FALSE, width = 0) +
  geom_point(data = t_opt_a_sum, aes(median, area, fill = factor(area, levels = rev(order$area))), shape = 21, color = "white", size = 3) +
  NULL

ggsave(paste0(home, "/figures/t_opt_ridges.pdf" ), width = 17, height = 17, units = "cm")

# Any patterns in the residuals (area spec and global)
t_opt_a_sum$global <- quantile(t_opt$temp, probs = c(0.5))

t_opt_a_sum$diff <- t_opt_a_sum$median - t_opt_a_sum$global

order2 <- VBG %>%
  ungroup() %>%
  group_by(area) %>%
  summarise(mean_temp = mean(temp)) %>%
  arrange(desc(mean_temp))

t_opt_a_sum <- left_join(t_opt_a_sum, order2)

diff <- ggplot(t_opt_a_sum, aes(mean_temp, diff, label = area)) +
  geom_point(color = "gray10") + 
  geom_errorbar(aes(ymin = lwr-global, ymax = upr-global), width = 0, alpha = 0.4) +
  labs(x = "Mean site temperature (°C)", y = "Site-specific - Global optimum") +
  geom_hline(yintercept = 0, alpha = 0.5, color = "gray30", linetype = 2) +
  NULL
  
(ridge & coord_flip() & guides(color = "none")) + diff + plot_layout(widths = c(1, 0.6)) + plot_annotation(tag_levels = "A")

ggsave(paste0(home, "/figures/t_opt_ridges.pdf" ), width = 17, height = 8, units = "cm", device = cairo_pdf)
```

### Plot area specific predictions

```{r}
#| cache: false
# Compare with area-specific sharpe scool!
area_pred_brms <- dat %>%
  group_by(area) %>% 
  data_grid(temp = seq_range(temp, n = 100)) %>% 
  ungroup() %>% 
  mutate(bk = 8.62e-05,
         tref = 8 + 273.15) %>% 
  add_epred_draws(fitbs)

# Plot area specific predictions as facets
p1 <- ggplot(area_pred_brms, aes(temp, y = .epred, color = factor(area, order$area),
                           fill = factor(area, order$area))) +
  stat_lineribbon(.width = c(0.95), alpha = 0.4, color = NA) +
  stat_lineribbon(.width = c(0)) + 
  geom_point(data = dat, aes(temp, rate, color = factor(area, order$area)), size = 0.6) +
  scale_color_manual(values = colors, name = "Site") +
  scale_fill_manual(values = colors, name = "Site") +
  scale_linetype_manual(values = 2) +
  facet_wrap(~factor(area, rev(order$area))) +
  guides(fill = "none",
         color = guide_legend(nrow = 1, reverse = TRUE, title.position = "top", title.hjust = 0.5,
                              override.aes = list(size = 1))) +
  theme(axis.title.y = ggtext::element_markdown(),
        legend.position = "bottom",
        legend.direction = "horizontal") +
  labs(x = "Temperature (°C)",
       y = "von Bertalanffy growth coefficient (*k*)",
       linetype = "")

p1

ggsave(paste0(home, "/figures/supp/sharpe_school_bayes_ci_facet.pdf" ), width = 17, height = 17, units = "cm")
ggsave(paste0(home, "/figures/for-talks/sharpe_school_bayes_ci_facet.pdf" ), width = 14, height = 14, units = "cm")
```

